---
title: "Hands-on_Exercise 1"
author: "YIN HANG"
date: "`r Sys.Date()`"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed. If they are, they will be launched into R.

```{r}
pacman::p_load(tidyverse, patchwork, 
               ggthemes, hrbrthemes,
               ggrepel)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# ggplot2

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```

# Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
```

# Plotting a geom_dotplot

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL) 
```

# Plotting a geom_histogram

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

Stat_summary() Method

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=4)               
```

Add Best Fit Curve on Scatterplot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method =lm,
              size=0.5)
```

Small Multiple Plots with Facets

facet_wrap()

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

facet_grid()

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
    facet_grid(GENDER ~ CLASS)
```

Coordinates

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

Work with X, Y axis limit

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

Bar Chart with Sorted Frequency

```{r}
ggplot(data=exam_data,
       aes(x=reorder(RACE,RACE,
                function(x)-length(x))))+
  geom_bar() +
  ylim(0,220) +
  geom_text(stat="count", 
      aes(label=paste0(..count.., ", ", 
      round(..count../sum(..count..)*100,
            1), "%")),
      vjust=-1) +
  xlab("Race") +
  ylab("No. of\nPupils") +
  theme(axis.title.y=element_text(angle = 0))
```

Mean and Median Line on Histogram Plot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  geom_vline(aes(xintercept=mean(MATHS,
                                 na.rm=T)),
             color="red", 
             linetype="dashed", 
             size=1) +
  geom_vline(aes(xintercept=median(MATHS,
                                  na.rm=T)),
             color="grey30",
             linetype="dashed", 
             size=1)
```

Histogram with Background Chart for Comparison

```{r}
d <- exam_data   
d_bg <- d[, -3]  

ggplot(d, aes(x = ENGLISH, fill = GENDER)) +
  geom_histogram(data = d_bg, fill = "grey", alpha = .5) +
  geom_histogram(colour = "black") +
  facet_wrap(~ GENDER) +
  guides(fill = FALSE) +  
  theme_bw()
```

A within group scatterplot with reference lines

```{r}
ggplot(data=exam_data, 
       aes(x=MATHS, y=ENGLISH)) +
  geom_point() +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  geom_hline(yintercept=50,
             linetype="dashed",
             color="grey60",
             size=1) + 
  geom_vline(xintercept=50, 
             linetype="dashed",
             color="grey60",
             size=1)
```
