---
title: "Hands-on Exercise 6"
author: "YIN HANG"
format: html
date: "18 Feb 2023"
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
editor: visual
---

# Getting Started

Load the code chunk to check whether the following packages has been installed

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, tidyverse, readxl, knitr, data.table)
```

# Plot Calendar Heatmap

### Load the Data

For the purpose of this hands-on exercise, *eventlog.csv* file will be used. This data file consists of 199,999 rows of time-series cyber attack records by country.

import *eventlog.csv* file into R environment and called the data frame as *attacks*.

```{r}
attacks <- read_csv("data/eventlog.csv")
```

### Examining the data structure

*kable()* can be used to review the structure of the imported data frame.

```{r echo=FALSE}
kable(head(attacks))
```

There are three columns, namely *timestamp*, *source_country* and *tz*.

-   *timestamp* field stores date-time values in POSIXct format.
-   *source_country* field stores the source of the attack. It is in *ISO 3166-1 alpha-2* country code.\
-   *tz* field stores time zone of the source IP address.

### Data Preparation

Step 1: Prepare weekday (wkday) and hour of day (hour) fields

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
  }
```

Step 2: Derive the attacks tibble data frame

```{r}
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

`mutate()` of **dplyr** package is used to convert *wkday* and *hour* fields into **factor** so they'll be ordered when plotting

Table below shows the tidy tibble table after processing.

```{r}
kable(head(attacks))
```

### Building Calendar Heatmap

```{r}
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

### **Building Multiple Calendar Heatmaps**

Step 1: Deriving attack by country object

In order to identify the top 4 countries with the highest number of attacks, we need to:

-   count the number of attacks by country,

-   calculate the percent of attackes by country, and

-   save the results in a tibble data frame.

    ```{r}
    attacks_by_country <- count(
      attacks, source_country) %>%
      mutate(percent = percent(n/sum(n))) %>%
      arrange(desc(n))
    ```

Step 2: Preparing the tidy data frame

extract the attack records of the top 4 countries from *attacks* data frame and save the data in a new tibble data frame (i.e. *top4_attacks*).

```{r}
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```

### Plotting Multiple Calendar Heatmaps

Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package.

```{r fig.width = 12}
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## Cycle Plot

### Data Preparation

#### Step 1: Data Import

For the purpose of this hands-on exercise, *arrivals_by_air.xlsx* will be used.

The code chunk below imports *arrivals_by_air.xlsx* by using `read_excel()` of **readxl** package and save it as a tibble data frame called *air*.

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

#### Step 2: Deriving month and year fields

Next, two new fields called *month* and *year* are derived from *Month-Year* field.

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

#### Step 4: Extracting the target country

Next, the code chunk below is use to extract data for the target country (i.e. Vietnam)

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

#### Step 5: Computing year average arrivals by month

The code chunk below uses `group_by()` and `summarise()` of **dplyr** to compute year average arrivals by month.

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

### Plotting the cycle plot

The code chunk below is used to plot the cycle plot

```{r eval=FALSE}
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors")
```
