---
title: "Hands-on_Ex05"
author: "YIN HANG"
date: "11 Feb 2023"
---

# Visualising Correlation Matrices

## Install and Launch R Packages

```{r}
pacman::p_load(corrplot, tidyverse, ggstatsplot)
```

## Import Dataset --Wine 

```{r}
wine <- read_csv("data/wine_quality.csv")
```

## Build Correlation Matrix

***pairs()*** **method**

```{r}
pairs(wine[,1:11])
```

***ggcormat()*** **method**

only correlation variables are needed (column 1 to 11)

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11)
```

Give more aesthetic argument and Control visual attributes: font size, X-Axis and Y-Axis

```{r}
ggstatsplot::ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black",
                         hc.order = TRUE,
                         tl.cex =10),
  title = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)

ggplot.component = list(
  theme(text = element_text(size = 5),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8)))
```

## Building multiple plots

```{r}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

***corrplot()*** **package**

Using cor() of R Stats to compute the correlation matrix of wine data frame

```{r}
wine.cor <- cor(wine[,1:11])
corrplot(wine.cor)
```

Change visual geometrics

```{r}
corrplot(wine.cor,
         method = "ellipse",
         type = "lower",
         diag = FALSE,
         tl.col = "black")
```

## Mixed layout with visual matrix and numerical matrix

```{r}
corrplot.mixed(wine.cor,
               lower = "ellipse",
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

## Corrgram with significant test

```{r}
wine.sig = cor.mtest(wine.cor, conf.level = 0.95)
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

### Reorder a corrgram

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```

### Reorder a correlation matrix with hclust

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```
