---
title: "Take-home_Ex03"
author: "YIN HANG"
date: "14 Feb 2023"
---

In this take-home exercise, we will look into the salient patterns of the resale prices of public housing property by residential towns and estates in Singapore.

# 1. R Preparation

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed. If they are, they will be launched into R.

```{r}
pacman::p_load(tidyverse, patchwork, 
               ggthemes, hrbrthemes,
               ggrepel,ggiraph)
```

## Importing the data

The code chunk below uses read_csv function to import the dataset: [*Resale flat princes based on registration date from Jan-2017 onwards*]{.underline}.

Data is from [Data.gov.sg](https://data.gov.sg/).

```{r}
property_data <- read_csv("data/resale-flat-prices-based-on-registration-date-from-jan-2017-onwards.csv")
head(property_data)
```

# Analytical Visualisation Selection

# Procedures for Analytical Visualisation

# Patterns Reveal
